\chapter{Installation of Software Libraries}
\label{appendix:installation_libraries}
This appendix will provide the details on how to install the software libraries used in this paper. The main programming language used in this paper is Julia \cite{Julia2017}, and can be installed by following the steps below:

\begin{enumerate}
    \item Go to the official Julia website: \url{https://julialang.org/downloads/}
    \item Download the latest version of Julia for your operating system (Windows, macOS, or Linux).
    \item Follow the installation instructions for your operating system.
    \item After installation, open the Julia REPL (Read-Eval-Print Loop) by running the `julia` command in your terminal or command prompt.
    \item To check if Julia is installed correctly, type `versioninfo()` in the Julia REPL and press Enter. You should see information about your Julia installation.
\end{enumerate}

Once succesffuly installed, next is to install the required libraries. The main libraries used in this paper are: QuranTree.jl \cite{asaad2021qurantree} and Yunir.jl \cite{al_ahmadgaid_b_asaad_yunir}, among others. The following steps will guide you on how to install these libraries:
\begin{enumerate}
    \item Open the Julia REPL.
    \item Type \verb|]| to enter the package manager mode.
    \item To install the QuranTree.jl library, type \verb|add QuranTree| and press Enter.
    \item To install the Yunir.jl library, type \verb|add Yunir| and press Enter.
    \item To install the other libraries, type \verb|add <library_name>| for each library you want to install. For example, to install the DataFrames.jl library, type \verb|add DataFrames|.
\end{enumerate}
