\chapter*{Dedication}
\addcontentsline{toc}{chapter}{Dedication}
\begin{center}
    \txarb{إلى الرجل الذي عُرف لدى البعض كمخطط ومتحدث عام، ولدى الكثيرين كمؤرخ، ومصمم، ونجار، وسائق تريكسيكل، وصياد، ولكن قبل كل شيء كأب متواضع، والدنا الحبيب، غروّاس س. أسعد، نسأل الله أن يرزقك الفردوس الأعلى، وأن يجمعنا جميعًا. أفتقدك كثيرًا يا أبي، شكرًا لكل شيء.}\\

    \textit{To the man known to some as a planner and a public speaker, but known to many as a historian, designer, carpenter, tricycle driver, fisherman, and most especially a humble father, our dear papa, Garwas S. Asaad. May Allah \txarb{\fontspec{Scheherazade New} ﷿} bless you with the highest of} \arb{jannaT} \textit{, and may He unite us all there together. I miss you so much pa, magsukul ma kamemon.}\\
    \begin{minipage}{.6\linewidth}
        \centering
        \vspace{4cm}
        \arb[fullvoc]{wa`ibAdu 'l-ra.hmAni 'lla_dIna yam^sUna `alaY" 'l-'ar.di hawnaN wa-'i_dA xA.tabahumu\\
        'l-j---------------------Ahi---------------------lUna 
        q---------------------Al---------------------UA" sa---------------------lAm---------------------aN}\\
        \textit{The servants of the Lord of Mercy are those who walk humbly on the earth, and who, when aggressive people address them, reply, with words of Peace.}\\
        {\sc Qur'an 25:63}\\[2cm]
        \arb[fullvoc]{'afalA yatadabbarUna 'l-qur'Ana walawkAna min `indi.gayri 'l-lahi lawajadUA" fIhi "axtilAfaN ka_tIraN}\\
        \textit{Will they not think about this Qur'an? If it had been from anyone other than God, they would have found much inconsistency in it.}\\
        {\sc Qur'an 4:82}\\[1cm]
    \end{minipage}
\end{center}