\chapter{Literature Review}

As mentioned in the previous chapter, the earliest work in Qur'\=anic studies using computer was likely the work of Rashad Khalifa in 1968\footnote{\url{https://www.masjidtucson.org/quran/miracle/a_profound_miracle_sura68nun133.html}}, which led to one of his book entitled `The Computer Speaks: God's Message to the World' (\textit{see} \citeA{rashad1981}). While the work of Rashad started at studying the mystery letters in the beginning of some \textit{s\=urahs} \arb{sUr} (for example Qur'\=an 2:1, 3:1, 7:1, etc.), it quickly went on to cover what he calls other \textit{mathematical miracles}, all of which are covered in \citeA{rashad1981}. His findings led him to generalize the claim that God revealed His words through this mathematical patterns throughout the Qur'\=an, and that those verses that were off and did not conform to this discovered mathematical patterns led him to extensive investigation of the said verses, and concluded that those could be or surely be an insertion that should not have been in the Qur'\=an in the first place. There are two verses that were off according to Rashad, and he called these verses as \textit{false verses}\footnote{\url{https://submission.org/App24.html}}, these are the last two \arb[trans]{'ayAt} \arb{'ayAt} of \arb[trans]{sUraT 'l-tAwbaT} \arb{sUraT 'l-tAwbaT} or The Chapter of \textit{Repentance}. These two verses were removed in Rashad's Qur'\=an translation\footnote{\url{https://www.masjidtucson.org/quran/frames/}}. Rashad believed so much on his findings that he self-proclaimed himself to be a messenger\footnote{\url{https://www.masjidtucson.org/submission/faq/rashad_khalifa_summary.html}} with this new findings and that the Qur'\=an nowadays should conform to his found mathematical patterns. This self-proclamation led to his assassination.

Fast forward to 20th century, among the pioneers to creating a stemmer systen for the Qur'\=an is the work of \citeA{thabet2004}. A stemmer system is a system for trimming inflected words into its basic form, which grammatically mean its the root form. For example, in English language the root word for \textit{computational}, \textit{computer}, \textit{computation}, and \textit{computerize} is \textit{compute}. Therefore, from the root of the word forms different stems representing the different words. Hence, the idea of stemming is to trim these words into its basic form, so that it would be easy to do word clustering or grouping through word similarity. According to \citeA{thabet2004}, the rich morphology of the Qur'\=anic language or the Classical Arabic makes it even more difficult to do word stemming. Moving on, \citeA{thabet2005} builds on this stemming system, and used it for tokenization of the Qur'\=anic words. Tokenization is the process of listing all of the words in a sentence, and \citeA{thabet2005} makes use of \citeA{thabet2004} to further cleanse the noise from these tokens brought by the morphological variants of the Classical Arabic words. After cleansing the data, \citeA{thabet2005} makes use of a statistical methodology for clustering or grouping the chapters of the Qur'\=an, in particular the statistical methodology used is the Agglomerative Hierarchical Clustering based on the Euclidean distance of the adjusted word frequency of the \textit{s\=urah} \arb{sUraT}.

The work of \citeA{thabet2004} and \citeA{thabet2005} makes use of the Qur'\=an corpus transliterated to Roman letters and symbols. Indeed, with interest growing on studying the Qur'\=an from the lense of Data Analysis and Natural Language Processing, more work have been put in place into creating digital corpi of the Qur'\=an that captures the different aspects of its linguistic styles. Hence, a series of work by Sharaf and Atwell led to the following publications: \citeA{sharaf2009} studied knowledge representation of the Qur'\=anic verb valences using FrameNet frames, the output of which is a lexical database of the corpus of Qur'\=an verbs. Further, the work of \citeA{sharaf2012} came up with corpus for the annotations of the Qur'\=anic pronouns, the authors named it as QurAna. Building on this work, \citeA{sharaf2012b} came up with a corpus for studying Qur'\=anic relatedness based on the commentary of Ibn Kathir \arb{ibn ka_tir}, the authors named this corpus as QurSim.

Moving on, an unpublished work by \citeA{nassourou2011} used a Support Vector Machine (SVM) to predict the classification of the place of revelation of the Qur'\=anic \textit{s\=urahs} \arb{sUr}. The idea was to use \textit{s\=urahs} \arb{sUr} with well attested place of revalation as the training set, and then train an SVM to predict the remainings of the chapters. Furthermore, the work of \citeA{shahzadi2012} developed a simple classifier based on the frequency distribution of words in \arb[trans]{'ayAt} \arb{'ayAt} and the corresponding words in \textit{s\=urahs} \arb{sUr}.

Moving on, the work of 